---
title: "BIS"
author: "Stefan Angrick"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BIS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `BIS` package package provides an `R` interface to data hosted by the [Bank for International Settlements](https://www.bis.org), specifically the [single-file data sets](https://www.bis.org/statistics/full_data_sets.htm) available on the BIS homepage.

## Install package

The package can be installed from GitHub.

```{r install, eval=FALSE, message=FALSE, warning=FALSE}
library(devtools)
install_github("stefanangrick/BIS")  # GitHub
```

## Example usage

To import data, first load the package:

```{r loading, message=FALSE, warning=FALSE}
library("BIS")
```

Next, run the `get_datasets()` function to obtain a list of available data sets:

```{r datasets, message=FALSE, warning=FALSE}
ds <- get_datasets()
head(ds, 20)
```

The `get_datasets()` function returns a [tibble](https://tibble.tidyverse.org/) data frame listing the available data sets. The column `url` can be used as input for the `get_bis()` function which downloads, parses and imports the corresponding data set.

To import monthly-frequency data on [central banks' policy rates](https://www.bis.org/statistics/cbpol.htm), run:

```{r rates, message=FALSE, warning=FALSE}
rates <- get_bis(ds$url[ds$id == "full_cbpol_m_csv"])
head(rates)
```

To plot the data using [ggplot2](https://ggplot2.tidyverse.org), run the following:

```{r plot, message=FALSE, warning=FALSE}
library("dplyr")
library("ggplot2")
library("zoo")

rates_plot <- subset(rates, ref_area %in% c("US", "XM", "JP", "GB", "CH", "CA"))
rates_plot <- mutate(rates_plot, date = as.Date(as.yearmon(date, format = "%Y-%m")))

ggplot(rates_plot, aes(date, obs_value, color = reference_area)) +
  geom_line(show.legend = FALSE) +
  facet_wrap(~reference_area) +
  labs(title = "Central bank policy rates",
       subtitle = "% per annum", x = NULL, y = NULL)
```

Note that BIS data sets come with a number of different time formats. The [zoo](https://cran.r-project.org/package=zoo) package (e.g. `as.yearmon()`) should be able to parse most formats.

## Large data sets

Large data sets (e.g. the Locational banking statistics and Debt securities statistics) may cause `get_bis()` to fail if the amount of available memory is insufficient for executing a required pivot operation. As a workaround, users may wish to set `auto_pivot = FALSE` when calling `get_bis()`, then subset the data and run `pivot_longer_bis()` manually.

```{r nopivot, eval=FALSE, message=FALSE, warning=FALSE}
options(timeout = 600)
lbs <- get_bis(ds$url[(ds$id == "full_lbs_d_pub_csv")], auto_pivot = FALSE)
lbs <- subset(lbs, l_parent_cty %in% c("US", "DE", "JP"))
lbs <- pivot_longer_bis(lbs)
```

## Retrieve individual data series

To retrieve individual data series instead of full data sets, consider using the BIS [SDMX RESTful API](https://stats.bis.org/api-doc/v1/#/). The [rsdmx R package](https://cran.r-project.org/package=rsdmx) is able to process SDMX data within R. The latest rsdmx [development version](https://github.com/opensdmx/rsdmx) contains a BIS connector that streamlines the process.

## Note

This package is in no way officially related to or endorsed by the [Bank for International Settlements](https://www.bis.org/). It's based on a fork of [CC0](https://cran.r-project.org/src/contrib/Archive/BIS/)-licensed [code by expersso](https://github.com/expersso/BIS). Please don't abuse the BIS's servers with unnecessary calls.
